---
title: "bzs surveys 2019"
author: "GG"
date: "October 30, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(dbplyr)
```
```{r}
bzs_preyfish <- read_csv("~/github/bzs_lionfish/preyfish_2019_raw.csv")
head(bzs_preyfish)
bzs_preyfish_clean <- filter(bzs_preyfish, size<20, density>0, depth!=45)
head(bzs_preyfish_clean)
```
```{r}
library(dplyr)
bzs_preyfish_final <- bzs_preyfish_clean %>%
  mutate(year=as.factor(year)) %>% mutate(depth=as.factor(depth))
head(bzs_preyfish_final)
```
```{r}
library(dbplyr)
mean_preyfish <- bzs_preyfish_final %>% group_by(year, site, depth, transect) %>% summarise(total_den = sum(density), total_bio = sum(biomass))
head(mean_preyfish)
```

```{r}
library(dbplyr)
mean_trophic <- bzs_preyfish_final %>% group_by(year, site, depth, transect, trophic_guild) %>% summarise(total_den = mean(density), total_bio = mean(biomass))
head(mean_trophic)
``` 
```{r}
biomass_trophicguild <- ggplot (data = mean_trophic, aes(x = year, y=total_bio, fill = trophic_guild)) + geom_bar(stat = "summary", fun.y = "mean", position = "fill", color = "black") + scale_fill_brewer(palette = "YlGnBu", labels = c("Herbivore", "Invertivore", "Macrocarnivore", "Omnivore", "Planktivore"), name = ("Trophic Guild")) + theme_classic() + labs(y="Percent Contribution to Biomass", x = "Survey Year") + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12)) + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14))
biomass_trophicguild
ggsave(filename = "biomass_trophicguild.eps", plot = biomass_trophicguild, path = "~/github/bzs_lionfish", device = "eps", width = 6,height = 4,dpi = 600)
```



```{r}

density_trophicguild <-ggplot(data = mean_trophic, aes(x = year, y=total_den, fill = trophic_guild)) + geom_bar(stat = "summary", fun.y = "mean", position = "fill", color = "black") + scale_fill_brewer(palette = "YlGnBu", labels = c("Herbivore", "Invertivore", "Macrocarnivore", "Omnivore", "Planktivore"), name = ("Trophic guild")) + theme_classic() + labs(y="Percent Contribution to Mean Density", x = "Survey Year") + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12)) + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14))
density_trophicguild
ggsave(filename = "density_trophicguild.eps",plot = density_trophicguild, path = "~/github/bzs_lionfish", device = "eps",width = 6,height = 4,dpi = 600)
```

```{r}
mean_biomass_depth <-ggplot(data = mean_preyfish, aes(x = depth, y=total_bio, fill = year)) + geom_bar(stat = "summary", fun.y = "mean", position = "dodge", color = "black") + labs(y = "Mean Biomass (g)", x = "Depth (m)") + stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge") + scale_fill_brewer(palette = "Paired", labels = c("2014", "2019"), name = ("Survey Year")) +theme_classic() + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12)) + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14))
mean_biomass_depth
ggsave(filename = "mean_biomass_depth.eps",plot = mean_biomass_depth, path = "~/github/bzs_lionfish", device = "eps",width = 6,height = 4,dpi = 600)
```

```{r}
mean_density_depth <-ggplot(data = mean_preyfish, aes(x = depth, y=total_den, fill = year)) + geom_bar(stat = "summary", fun.y = "mean", position = "dodge", color = "black") + labs(y = "Mean Density (#/30m^2)", x = "Depth (m)") + stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge") + scale_fill_brewer(palette = "Paired", labels = c("2014", "2019"), name = ("Survey Year")) +theme_classic() + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12)) + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14))
mean_density_depth
ggsave(filename = "mean_density_depth.eps",plot = mean_density_depth, path = "~/github/bzs_lionfish", device = "eps",width = 6,height = 4,dpi = 600)
```
```{r}
trophic_density_depth <-ggplot(data = mean_trophic, aes(x = depth, y=total_den, fill = trophic_guild)) + geom_bar(stat = "summary", fun.y = "mean", position = "fill", color = "black") + labs(y = "Mean % Contribution", x = "Depth (m)") + scale_fill_brewer(palette = "YlGnBu", labels = c("Herbivore", "Invertivore", "Macrocarnivore", "Omnivore", "Planktivore"), name = ("Trophic guild")) +theme_classic() + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12)) + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14))
trophic_density_depth
ggsave(filename = "mean_density_depth.eps",plot = mean_density_depth, path = "~/github/bzs_lionfish", device = "eps",width = 6,height = 4,dpi = 600)
```
```{r}
trophic_density_depth_year <-ggplot(data = mean_trophic, aes(x = depth, y=total_den, fill = trophic_guild)) + geom_bar(stat = "summary", fun.y = "mean", position = "stack", color = "black") + labs(y = "Mean Density", x = "Depth (m)") + scale_fill_brewer(palette = "YlGnBu", labels = c("Herbivore", "Invertivore", "Macrocarnivore", "Omnivore", "Planktivore"), name = ("Trophic guild")) +theme_bw() + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12)) + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) + facet_grid(year~.)
trophic_density_depth_year
```
```{r}
density_depth_year <-ggplot(data = mean_preyfish, aes(x = depth, y=total_den, fill = depth)) + geom_bar(stat = "summary", fun.y = "mean", position = "stack", color = "black") + labs(y = "Mean Density", x = "Depth (m)") + stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge") + scale_fill_brewer(palette = "Blues") +theme_bw() + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12)) + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) + facet_grid(year~.)
density_depth_year
```

```{r}
mean_biomass_overall <-ggplot(data = mean_preyfish, aes(x = year, y=total_bio, fill = year)) + geom_bar(stat = "summary", fun.y = "mean", position = "dodge", color = "black") + labs(y = "Mean Biomass (g)", x = "Year") + stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge") + scale_fill_brewer(palette = "Paired") +theme_classic() + theme(legend.position = "none")+theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14))
mean_biomass_overall
ggsave(filename = "mean_biomass_overall.eps",plot = mean_biomass_overall, path = "~/github/bzs_lionfish", device = "eps",width = 6,height = 4,dpi = 600)
```

```{r}
mean_density_overall <-ggplot(data = mean_preyfish, aes(x = year, y=total_den, fill = year)) + geom_bar(stat = "summary", fun.y = "mean", position = "dodge", color = "black") + labs(y = "Mean Density (#fish/30m^2)", x = "Year") + stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge") + scale_fill_brewer(palette = "Paired") +theme_classic() + theme(legend.position = "none")+theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14))
mean_density_overall
ggsave(filename = "mean_density_overall.eps",plot = mean_density_overall, path = "~/github/bzs_lionfish", device = "eps",width = 6,height = 4,dpi = 600)
```





```{r}
biomass_site_depth_year <-ggplot(mean_preyfish, aes(year, total_bio)) + geom_bar(aes(fill = depth, alpha = year), position = "dodge", stat = "summary", fun.y = "mean", color = "black") + labs(y = "Mean biomass (g)", x = "Site") + stat_summary(geom = "errorbar", fun.data = mean_se)+
  scale_alpha_manual(values = c(0.3, 1)) + scale_fill_brewer(palette = "YlGnBu") +
  facet_grid(depth~site, scales = "free")  + theme_bw() + theme(axis.text.x = element_text(angle = 45, size = 12, hjust = 1)) + theme(legend.position = "none") + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14, face = "bold"))
biomass_site_depth_year
```
```{r}
preyfishsubset <- mean_preyfish %>% subset(depth == 60)
head(preyfishsubset)
meso_prey_bio <-ggplot(data = preyfishsubset, aes(x = site, y=total_bio, fill = year)) + geom_bar(stat = "summary", fun.y = "mean", position = "dodge", color = "black") + labs(y = "Mean biomass (g)", x = "Site") + stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge") + scale_fill_brewer(palette = "Paired", labels = c("2014", "2019"), name = ("Survey year")) +theme_classic() + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12)) + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) + theme(axis.text.x = element_text(angle = 20, hjust = 1))
meso_prey_bio
```
```{r}
meso_prey_den <-ggplot(data = preyfishsubset, aes(x = site, y=total_den, fill = year)) + geom_bar(stat = "summary", fun.y = "mean", position = "dodge", color = "black") + labs(y = "Mean density (#/30 m^2)", x = "Site") + stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge") + scale_fill_brewer(palette = "Paired", labels = c("2014", "2019"), name = ("Survey year")) +theme_classic() + theme(legend.title = element_text(size = 14), legend.text = element_text(size = 12)) + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14))+ theme(axis.text.x = element_text(angle = 20, hjust = 1))
meso_prey_den
```
```{r}
shapiro.test(mean_preyfish$area)
```
```{r}
shapiro.test(sqrt(past_size$area))
```
```{r}
library(WRS2)
```
```{r}
GROUP<- factor(preyfish, c(1,2),labels=c('survey','depth'))
WC1 <- wilcox.test(biomass~GROUP, data = preyfish)
WC1
#doesnt work with both survey and depth; says that grouping factor must have exactly 2 levels; doesnt seem like it is the right thing to use
```

```{r}
install.packages("multcompView")
library(multcompView)
library(car)
anova1 <- aov((total_bio) ~ depth + year + year/depth, data = mean_preyfish)
shapiro.test(sqrt(sqrt(sqrt((mean_preyfish$total_den))))) #not met, p=2.2e-16
leveneTest(mean_preyfish$total_den) #not met, p=1.103e-10
summary.aov(anova1)
TukeyHSD(anova1)
```
```{r}
fit3 <- glm((total_bio)~year + depth + year*depth, data=mean_preyfish)
summary(fit3)
```

Results
1.) Year and depth impact the biomass of prey fish 
2.) There are sigificantly more fish in 2019 than in 2014
3.) There are signifcantly more fish at 60m then at any other depth 
4.) 2014-2019 at 10m and 60m prey fish biomass signifcantly increased

```{r}
#subset based on trophic guild then compare it to survey
#Subset to 60m then biomass by trophic guild then depth 
trophicsubset <- preyfish %>% subset(depth == 60)
anova2 <- aov(sqrt(density) ~ trophic_guild, data = trophicsubset)
shapiro.test(anova2$residuals) #not met, p=2.2e-16
leveneTest(anova2) #not met, p=2.2e-16
summary.aov(anova2)
TukeyHSD(anova2)
```
Results
1.) Plantivorous fish density is significantly greater than all other trophic guilds at 60m
2.) Reference gretchens model on plankton abundance and upwelling 

```{r}
trophicPsubset <- preyfish %>% subset(trophic_guild == "Planktivore")
anova3 <- aov(sqrt(density) ~ survey* depth, data = trophicPsubset)
shapiro.test(anova3$residuals) #not met, p=0.01774
leveneTest(anova3) #not met, p=2.353e-05
summary.aov(anova3)
TukeyHSD(anova3)
```

Results
1.) Depth and survey year impacted planktivorous density 
2.) Plantivorous fish significantly increased at 60m across survey year
3.) Interesting- are lionfish filling a predator role niche that is absent on the reefs? Fish community is not drastically changing with lionfish density staying fairly the same 

```{r}
trophicIsubset <- preyfish %>% subset(trophic_guild == "Invertivore")
anova4 <- aov(sqrt(density) ~ survey* depth, data = trophicIsubset)
shapiro.test(anova4$residuals) #not met, p=2.2e-16
leveneTest(anova4) #met, p=0.1306
summary.aov(anova4)
```
Results
1.) Depth and survey year did not impact invertivore density 

```{r}
trophicOsubset <- preyfish %>% subset(trophic_guild == "Omnivore")
anova5 <- aov(sqrt(density) ~ survey* depth, data = trophicOsubset)
shapiro.test(anova5$residuals) #not met, p=2.2e-16
leveneTest(anova5) #not met, p=1.084e-06
summary.aov(anova5)
TukeyHSD(anova5)
```
Results
1.) Depth and survey year impacted omnivorous density 
3.) Omnivorous fish significantly decreased at 10m across survey year
4.) How much is overfishing in the shallow areas affecting fish biomass- complete loss of omnivorous fish species- the ones that we are catching?

```{r}
trophicMsubset <- preyfish %>% subset(trophic_guild == "Macrocarnivore")
anova6 <- aov(sqrt(density) ~ survey * depth, data = trophicMsubset)
shapiro.test(anova6$residuals) #not met, p=7.37e-05
leveneTest(anova6) #met, p=0.2015
summary.aov(anova6)
```
Results
1.) Depth and survey year did not impact macrocarnivore density 

```{r}
anova7 <- aov(sqrt(density) ~ survey * depth, data = preyfish)
shapiro.test(anova7$residuals) #not met, p=2.2e-16
leveneTest(anova7) #Not met, p=3.2e-15
summary.aov(anova7)
TukeyHSD(anova7)
```
Results
1.) Depth and year impact the trophic guild composition and the density of prey fish
2.) Plantivorous fish density is significantly greater than all other trophic guilds

```{r}
trophicPsubset <- preyfish %>% subset(trophic_guild == "Planktivore")
anova8 <- aov(log(biomass) ~ survey* depth, data = trophicPsubset)
shapiro.test(anova8$residuals) #met, p=0.2949
leveneTest(anova8) #met, p=0.2152
summary.aov(anova8) #depth and survey were sigificant independently
TukeyHSD(anova8)
```

Results
1.) Depth and survey year impacted planktivorous biomass 

```{r}
trophicIsubset <- preyfish %>% subset(trophic_guild == "Invertivore")
anova9 <- aov(log(biomass) ~ survey* depth, data = trophicIsubset)
shapiro.test(anova9$residuals) #not met, p=2.0009e-5
leveneTest(anova9) #not met, p=0.0005574
summary.aov(anova9)
```
Results
1.) Depth and survey year did not impact invertivore biomass

```{r}
trophicOsubset <- preyfish %>% subset(trophic_guild == "Omnivore")
anova10 <- aov(log(biomass) ~ survey* depth, data = trophicOsubset)
shapiro.test(anova10$residuals) #not met, p=1.232e-5
leveneTest(anova10) #met, p=0.7257
summary.aov(anova10)
TukeyHSD(anova10)
```
Results
1.) Depth impacted omnivorous biomass


```{r}
trophicMsubset <- preyfish %>% subset(trophic_guild == "Macrocarnivore")
anova11 <- aov(log(biomass) ~ survey * depth, data = trophicMsubset)
shapiro.test(anova11$residuals) #met, p=0.1201
leveneTest(anova11) #met, p=0.4255
summary.aov(anova11)
```

Results
1.) Depth and survey year did not impact macrocarnivore biomass

```{r}
sixtysubset <- preyfish %>% subset(depth > 59)
anova12 <- aov(log(biomass) ~ site *survey, data = sixtysubset)
shapiro.test(anova12$residuals) #not met, p=0.0001515
leveneTest(anova12) #not met, p=1.973e-5
summary.aov(anova12)
```
```{r}
sixtysubset <- preyfish %>% subset(depth > 59)
anova13 <- aov(log(density) ~ site *survey, data = sixtysubset)
shapiro.test(anova13$residuals) #not met, p=3.692e-10
leveneTest(anova13) #not met, p=0.00013
summary.aov(anova13) 
TukeyHSD(anova13)
```


